### 2013-02-12 doc/icml2013/algo.tex
Spruced up algo section and added details for the regression and tensor power method.

The goal is to recover the parameters $(\pi, B)$.
What should the moments be?

OUTLINE

\subsection{Recovering $B$ and $\pi$ from $M_2,M_3$}

If we knew the following quantities, then we could apply tensor factorization
\cite{AnandkumarGeHsu2012}.

Having recovered the compound parameters $M_2$ and $M_3$, we must now solve for the
actual parameters $\{\beta_k\}$ and $\{\pi_k\}$, which are the
eigenvectors and eigenvalues of the symmetric moments respectively,
\begin{eqnarray*}
  M_2 &=&  \sum_{k=1}^{K} \pi_k \beta_k^{\otimes 2}\\
  M_3 &=& \sum_{k=1}^{K} \pi_k \beta_k^{\otimes 3}.
\end{eqnarray*}

@@

Tensor power method

In this section, we will outline some shortcomings of the standard
method of moments formulation for the mixture of linear regressions
model. We then describe how we can work around this limitation and
reduce our problem to a known one, which can be solved using symmetric
tensor decomposition. Finally, we will state some implications this
reduction has regarding identifiability.

We want to estimate from low-order moments.

\subsection{Recovering $M_2$ and $M_3$}

Conventionally, when using the method of moments, the moments of
observed variables are expressed algebraically in terms of the model
parameters. Using these equations, we attempt to solve for the model
parameters using empirical estimates of the moments.

In the case of the mixture of linear regressions, we have two observed
quantities, $x$ and $y$. Unfortunately, because $y$ is a scalar
quantity, we have only one equation for every moment of $y$, and with
this naive approach, would have to consider $O( (K+1) D)$ moments to
recover $B \in \Re{K \times D}$ and $\pi \in \Re^K$.

To work around this problem, we will use regression on the conditional
moments, $\E[y \mid x]$ to recover the expected moments of the regressors,
$M_2$ and $M_3$. Recovering the $\beta_k$ and $\pi_k$ given these
symmetric moments has been addressed in prior work
\citep{AnandkumarHsuKakade2012, AnandkumarGeHsu2012}.  This algorithm as
been summarized in \algorithmref{algo:spectral-experts}.

@@

Let $\epsilon \sim \normal{0}{\sigma^2}$ be independent Gaussian noise.
we have,
\begin{eqnarray*}
  %y &=& \sum_{k=1}^{K} \delta_{k} \beta_k^T x + \epsilon \\
  \E[ y | x ] &=& M_1^T x \\
  \E[ y^2 | x ] &=& \sum_{k=1}^{K} \pi_{k} (\beta_k^T x)^2 + \E[ \epsilon^2 ] \\
  &=& M_2 \odot x^{\otimes 2} + \sigma^2 \\
  \E[ y^3 | x ] &=& \sum_{k=1}^{K} \pi_{k} (\beta_k^T x)^3
            + 3 \E[\epsilon^2] (\sum_{k=1}^{K} \pi_{k} \beta_k^T) x \\
  &=& M_3 \odot x^{\otimes 3} + 3 \sigma^2 M_1.
\end{eqnarray*}

% Details about low rank regression.
%The symmetric structure of $M_2$ and $M_3$ implies that there are
%$\frac{d (d-1)}{2}$ and $\frac{d (d-1) (d-2)}{6}$ degrees of freedom
%respectively.
However, the moments are further constrained to have
a rank of $K$. To exploit this property, we will use trace norm
regularization with a proximal subgradient descent
algorithm\citationneeded to recover $M_2$,
\begin{align}
  %\hat M_2 &= \arg\min_{M_2} \frac{1}{2 N} \sum_{i=1}^N ((\xn{i})\tp{2} \odot M_2 - (\yn{i})^2)^2 + \lambda \|M\|_*.
  \hat M_2 &= \arg\min_{M_2} \frac{1}{2 N} \sum_{(x,y) \in D} (x\tp{2} \odot M_2 - y^2)^2 + \lambda \|M\|_*.
\end{align}

The analogue of trace norm regularization for higher order tensors
corresponds to the sum of the trace norms of the mode-k unfolding of the
tensor \cite{Tomioka2011}, $X_{(k)}$, is a $d_k \times (\prod_{k' \neq
k} d_{k'})$ matrix obtained by concatenating the entries for all
dimensions other than $k$. For example, the 1-mode unfolding of a 3rd
order tensor has entries, $X_{(1)}^{i_1, (i_2, i_3)} = X_{i_1, i_2,
i_3}$. Thus, we employ a proximal subgradient descent algorithm to solve
the following optimization problem to recovery $M_3$.
\begin{eqnarray*}
  M_3 &=& \arg\min_{M} \half \sum_{i=1}^N ((\xn{i})\tp{3} \odot M - (\yn{i})^3)^2 
          + \frac{\lambda}{3} \sum_{k=1}^{3} \|M_{(k)}\|_*.
\end{eqnarray*}

In \sectionref{sec:theory}, we will show that recovery of matrices so
constrained requires only $O(\frac{\sqrt{K} d}{\epsilon})$ samples as
opposed to $O(\frac{d^3}{\epsilon})$ samples to recover moments within
$\epsilon$ from their true values.

@@

The model is in general identifiable, except for a pathological case
concerning how data points are placed \citationneeded.

Our approach employs regression on $x\tp{2}$ and $x\tp{3}$, and thus
requires $\mathcal{X} \otimes \mathcal{X}$ and $\mathcal{X} \otimes
\mathcal{X} \otimes \mathcal{X}$ to be linearly independent modulo
symmetry. We call these condition {\em quadratic} and {\em cubic}
independence.

@@

{\todo Note: we haven't shown identifiability from all moments. What
does this mean?}

>>>>

In this section, we will outline an algorithm to recover the parameters
$\pi$ and $B$ of the mixture of linear regressions model.

% Motivate how we can recover $B$ from regression by lower order moments
Our goal is to recover the parameters $(\pi, B)$, ideally from lower
order moments. Conventionally, when using the method of moments, the
moments of observed variables are expressed algebraically in terms of
the model parameters. Using these equations, we attempt to solve for the
model parameters using empirical estimates of the moments.

Let us consider the response variable $y$, 
\begin{align}
  y &=& \innerp{\beta_h}{x} + \epsilon,
\end{align}
where $\epsilon \sim \normal{0}{\sigma^2}$ is the observation noise.  It
is natural to consider regression on the data $\{\xni,
\yni\}_{i=1}^{N}$, but it would only help us find the mean regression
coefficients, $M_1 = \E[\beta_h] = \sum_h \pi_h \beta_h$, which are insufficient to
identify the model. Let us consider some more low powers of $y$, noting
that taking the expectation gives us the corresponding low-order
moments,
\begin{align}
  y^2 &= \innerp{\beta_h\tp{2}}{x\tp{2}} + \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} + \epsilon^2 \label{eq:y2} \\ 
  y^3 &= \innerp{\beta_h\tp{3}}{x\tp{3}} + 3 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} \notag \\
  &+ 3 \epsilon^2 \innerp{\beta_h}{x} + \epsilon^3. \label{eq:y3} 
\end{align}
Regression with the second order moments would give us $M_2 = \E[\beta_h
\tp{2}] = \sum_h \pi_h \beta_h\tp{2}$ which can identify $\pi$ and the
subspace spanned the $\beta_h$, but not the $\beta_h$ themselves. To see
this, note that the eigen-decomposition of $M_2$ is not unique.
\todo{explain better}.

This brings us to regression with the third order moments, giving us the
tensor $M_3 = \E[\beta_h \tp{3}] = \sum_h \pi_h \beta_h\tp{3}$.
\citet{AnandkumarGeHsu} show that these three moments suffice to
identify $\pi$ and $B$ and provide an efficient algorithm to recover the
parameters.

@@

% Describe the rest of the algorithm.
This description completes a sketch of the algorithm, described in
\algorithmref{algo:spectral-experts}. Going ahead, we have yet to show
that the regression is well-behaved which we will do in
\sectionref{sec:regression}. This is of concern because the regression
problem we have has variance introduced from component selection,
independent of any observation noise. We will show that we can indeed
efficiently recover $M_2$ and $M_3$ using ideas from low-rank
regression. Finally, we will outline the tensor power method to recover
$B$ and $\pi$ given these two quantities, $M_2$ and $M_3$ in
\sectionref{sec:tensor-power}.

\begin{theorem}
  Recovery analysis of mixture of linear regressions.
\end{theorem}
\begin{proof}
\end{proof}

@@

We will conclude this section by
discussing some identifiability concerns for our algorithm.

@@

The mixture of linear regressions model is identifiable subject to
permutations of the components\citationneeded. However, our approach
employs regression on $x\tp{2}$ and $x\tp{3}$, and thus requires
$\mathcal{X} \otimes \mathcal{X}$ and $\mathcal{X} \otimes \mathcal{X}
\otimes \mathcal{X}$ to be linearly independent modulo symmetry. We call
these condition {\em quadratic} and {\em cubic} independence.

@@

% \subsection{Recovering $M_2$ and $M_3$}

% In the case of the mixture of linear regressions, we have two observed
% quantities, $x$ and $y$. Unfortunately, because $y$ is a scalar
% quantity, we have only one equation for every moment of $y$, and with
% this naive approach, would have to consider $O( (K+1) D)$ moments to
% recover $B \in \Re{K \times D}$ and $\pi \in \Re^K$.

### 2013-02-12 doc/icml2013/intro.tex
Spruced up algo section and added details for the regression and tensor power method.

Latent variable models have routinely been used in probabilistic
modelling, (examples). Parameters for these models have been learnt
through a variety of methods, EM, variational methods, etc. The local
nature of these algorithms lead them to converge to local optima.
Recently, there has been an interest in moment of methods estimators
with polynomial sample and computational complexities. These estimators
are statistically consistent, i.e. given sufficient data they are
guaranteed to converge to the global optimum in an identifiable model,
thus sidestepping the local optima issue altogether.

% Discriminative models work better in practice
% Discriminative models with latent variables add expressive power (examples)
This approach has been applied to a variety of generative models, e.g.
hidden Markov models \cite{AnandkumarHsuKakade2012}, probabilistic
context free grammars \cite{HsuKakadeLiang2012}, etc. However, in
practice it has been observed that discriminative models work better in
terms of parameter recovery\citationneeded, particularly in the large
sample regime. \todo{Examples! CRFs, etc.}

% This paper: first step in establishing provably correct estimation
% with simple mixture of linear regression
In this work, we provide a first step in establishing provably correct
estimation for such discriminative models by proposing an algorithm for
the simple mixture of linear regressions. The mixture of linear
regressions model \todo{models ?}. We describe the model formally in more
detail in \sectionref{sec:model}.

% Previous work: moment equations solvable by eigendecomposition or
% tensor factorization
While method of moment estimators were \todo{described by Fisher(?)},
they required as many moments as there were parameters, and hence deemed
impractical. Recently, \citet{AnandkumarHsuKakade2012} showed how using
eigendecomposition, the moment equations could be solved using just the
first three moments of data.

% Challenge: E[X^m Y^n] doesn't work
In the discriminative setting, we are trying to learn the parameters for
a model for $y | X$. However, as the response variables are just
scalars, it is not possible to recover the moments simply from $\E[ X^m
y^n ].$ At the same time, knowing the $X$ should give us more
information than the scalar $y$.

% Approach: Derive moment functions based on linear regression on Y^2
% and Y^3 Algorithm: trace norm regularization to recover tensor
% + tensor factorization
We propose an algorithm in \sectionref{sec:algo} that derives moment
functions based on linear regression on the moments of $y$, i.e.
$\E[y^2]$ and $\E[y^3]$. We exploit the low rank properties of the
second and third order moments to efficiently recover parameters, and
then recover the individual regressors, $\beta_k$, using the tensor
factorization method \cite{AnandkumarGeHsu2012}.

@@

>>>>

Latent variable models have routinely been used in probabilistic
modelling, (examples). Parameters for these models have been learnt
through a variety of methods, EM, variational methods, etc. The local
nature of these algorithms lead them to converge to local optima.
Recently, there has been an interest in method of moments estimators for
generative models such as the hidden Markov
model\cite{AnandkumarHsuKakade2012} or probabilistic context free
grammars\cite{HsuKakadeLiang2012}.These estimators feature polynomial
sample and computational complexities while being statistically
consistent, i.e. given sufficient data they are guaranteed to converge
to the global optimum in an identifiable model, thus sidestepping the
local optima issue altogether.

% Discriminative models work better in practice Discriminative models
% with latent variables add expressive power (examples)
In practice it has been observed that discriminative models work better
in terms of parameter recovery\citationneeded, particularly in the large
sample regime. \todo{Examples! CRFs, etc.}
% This paper: first step in establishing provably correct estimation
% with simple mixture of linear regression
In this work, we provide a first step in establishing provably correct
estimation for such discriminative models by proposing an algorithm for
the simple mixture of linear regressions. We describe the model formally
in more detail in \sectionref{sec:model}.

% Previous work: moment equations solvable by eigendecomposition or
% tensor factorization
%While method of moment estimators were \todo{described by Fisher(?)},
%they required as many moments as there were parameters, and hence deemed
%impractical. Recently, \citet{AnandkumarHsuKakade2012} showed how using
%eigendecomposition, the moment equations could be solved using just the
%first three moments of data.

% Challenge: E[X^m Y^n] doesn't work
%In the discriminative setting, we are trying to learn the parameters for
%a model for $y | X$. However, as the response variables are just
%scalars, it is not possible to recover the moments simply from $\E[ X^m
%y^n ].$ At the same time, knowing the $X$ should give us more
%information than the scalar $y$.

% Approach: Derive moment functions based on linear regression on Y^2
% and Y^3 Algorithm: trace norm regularization to recover tensor
% + tensor factorization
We propose an algorithm in \sectionref{sec:algo} that derives moment
functions based on linear regression on the conditional moments of $y$,
i.e. $\E[y^2 | X]$ and $\E[y^3 | X]$. We exploit the low rank properties
of the second and third order moments to efficiently recover parameters,
and then recover the individual regressors, $\beta_k$, using the tensor
factorization method \cite{AnandkumarGeHsu2012}.

@@

\subsection{Notation}

%We have already defined $y$ to be the response variable and $x \in
%\Re^d$ to be the covariates it depends on. Note that the matrix of
%regressors $B$ is a $d \times K$ matrix, where the $k$-th column is
%$\beta_k$.

We will use the notation $x\tp{p}$ to describe the $p$-th order tensor
formed by taking the outer product of $x$; i.e. $x\tp{p}_{i_1 i_2 \ldots
i_p} = x_{i_1} x_{i_2} \cdots x_{i_p}$. We will use
$\innerp{\cdot}{\cdot}$ to denote the general dot product between $p$-th
order tensors, i.e. $\innerp{X}{Y} = \sum_{i_1, i_2, \ldots i_p} X_{i_1,
i_2, \ldots i_p} Y_{i_1, i_2, \ldots i_p}.$ Similar to how $x^T A$ is
the projection of a vector $x$ onto the row space of a matrix $A$, when
$X \in \Re^{d^p}$ and $Y \in \Re^{d^{p+q}}$,  with $p, q > 0$,
$\innerp{X}{Y} \in \Re^{d^q}$ is the projection of $X$ onto the first
$p$ dimensions of $Y$, i.e. $(\innerp{X}{Y})_{i_1, \ldots, i_q}
= \sum_{j_1, \ldots, j_p} X_{j_1, \ldots, j_p} Y_{j_1, \ldots, j_p, i_1,
\ldots, i_q}$.

The expected moments of the parameters $\beta_1, \dots, \beta_k$ play an
important role in our algorithm and we will use $M_p = \E[\beta\tp{p}]
= \sum_{k=1}^{K} \pi_k \beta_k^{\otimes p}$ to denote them.

### 2013-02-12 doc/icml2013/model.tex
Spruced up algo section and added details for the regression and tensor power method.

The mixture of linear regressions model describes response variables $y
\in \Re$ dependent on covariates $x \in \Re^d$ through $K$ different
modes which occur with probability $\pi_1, \dots, \pi_K$ respectively.
Each mode is independently described by regression coefficients
$\beta_k$ and we consider a common variance term $\sigma^2$. Given $x$,
the generative process for the response variables is,
\begin{eqnarray*}
  h &\sim& \mult(\pi) \\
  y | h = k &\sim& \normal{ \beta_{k}^T x}{ \sigma^2 },
\end{eqnarray*}
where $h$ is a latent variable corresponding to the mode.

@@

>>>>

The mixture of linear regressions model\citep{VieleTong} describes
response variables $y \in \Re$ dependent on covariates $x \in \Re^d$
through $K$ different modes which occur with probability $\pi_1, \dots,
\pi_K$ respectively. Each mixture component is independently described
by regression coefficients $\beta_k$ and we consider a common variance
term $\sigma^2$. Given $x$, the generative process for the response
variables is,
\begin{eqnarray*}
  h &\sim& \mult(\pi) \\
  y \mid h &\sim& \normal{ \beta_{h}^T x}{ \sigma^2 },
\end{eqnarray*}
where $h$ is a latent variable corresponding to the mode.

@@

% Typically people use EM, but the maximum marginal likelihood is
% non-convex.
Typically, the parameters $\pi, B$ are learnt usig
expectation-maximization\cite{VieleTong}, but the maximum marginal
likelihood is non-convex. We provide several instances where this leads
to poor local optima in \sectionref{sec:evaluation}.

% Reference mixture of experts. 
% Acknowledge that mixing proportions % can't depend on $x$
The mixture of linear regressions model is closely related to the
mixture of experts model\citationneeded, wherein the components can be
SVMs, GPs, regression trees and other estimators. In practice, it is
desirable to have mixture proportions that are dependent on $x$. We have
not considered these extensions in our treatment of the model.
